---
title: "36-315 Homework 10, Spring 2018"
author: "Ross Hemmel (rhemmel), Joshua Huang (jzh1), Julie Kim (juliek1), Jinhee Lee (jinheel1)"
date: "Due Wednesday, April 18, 2018 (11:59pm) on Canvas"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---

#  Homework 10: An Interesting Duo - Networks and Time Series

***
***

# Problem 0

* "We used Google's R style guide."
* "We used Hadley Wickham's R style guide."

* "This submission is for the following group: Ross Hemmel (rhemmel), Joshua Huang (jzh1), Julie Kim (juliek1), Jinhee Lee (jinheel1)

***
***


#  Problem 1

a.

b.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
love_adjacency <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/love-actually/love_actually_adjacencies.csv")
```

c. `ggraph` is an extention of `ggplot2` that allows a user to create network graphs with `ggplot2`. `ggraph` uses the `theme_graph()` function to override the typical `ggplot2` theme. `facet_nodes()`, `facet_edges()` and `facet_graph()` allow the user to plot marginal graphs by separating the data by some chosen variable.

d. 

```{r, warning = FALSE, message = FALSE}
library(igraph)
library(ggraph)

actor_names <- love_adjacency[,1]
LA_graph <- graph_from_adjacency_matrix(as.dist(love_adjacency[,-1]))

ggraph(LA_graph, layout = "kk") +
  geom_edge_link() +
  geom_node_label(aes(label = actor_names), size = 1.75) +
  theme_graph() +
  labs(title = "\"Love Actually\" Character Interactions",
       caption = "fivethirtyeight.com")
```

e. 

```{r}
LA_graph_weighted <- graph_from_adjacency_matrix(as.matrix(love_adjacency[,-1]),
                                        weighted = TRUE)

ggraph(LA_graph_weighted, layout = "lgl") +
  geom_edge_link(aes(edge_width = weight)) +
  scale_edge_width(range = c(0.5, 3)) +
  geom_node_label(aes(label = actor_names), size = 1.75) +
  theme_graph() +
  labs(title = "Relative Shared Screen Time of Characters in \"Love Actually\"",
       edge_width = "Relative Time",
       caption = "fivethirtyeight.com")

```


***
***

#  Problem 2

a. 

```{r, warning = F, message = F}
library(tidyverse)

# Only use the June 2016 subset
bike <- read_csv("https://github.com/mateyneykov/315_code_data/raw/master/data/nyc-citi-bike-data-subset.csv")
#  Add start_date variable
bike <- mutate(bike,
               start_date = as.Date(starttime, format = "%m/%d/%Y"))

```

+  `bike` has 4359 rows and 17 columns
+  The earliest trip date is `r bike$start_date %>% min`, and the latest trip date is `r bike$start_date %>% max`.

b.

```{r, warning = FALSE, message = FALSE}
#  Summarize bike, creating a new data.frame that includes the number 
#  of trips taken on each day

trips_per_day <- bike %>%
  group_by(start_date) %>%
  summarize(n_trips = n())

#  Create a time series plot with the dates on the x-axis and the number of
#  trips per day on the y-axis

ggplot(trips_per_day, aes(x = start_date, y = n_trips)) + geom_line() + 
  scale_x_date(date_breaks = "1 week", date_labels = "%B %d") +
  labs(title = "CitiBike Trips per Day",
       x = "Day",
       y = "Count",
       caption = "https://www.citibikenyc.com/system-data") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

c. 

```{r, warning = FALSE, message = FALSE}
#  Summarize bike, creating a new data.frame that includes the number 
#  of trips taken on each day, split by usertype
trips_per_day_usertype <- bike %>%
  group_by(start_date, usertype) %>% summarize(n_trips = n())

#  Create a time series plot with the dates on the x-axis and the number of
#  trips per day on the y-axis, split by usertype

ggplot(trips_per_day_usertype, aes(x = start_date, y = n_trips, color = usertype)) + 
  geom_line() + 
  scale_x_date(date_breaks = "1 week", date_labels = "%B %d") +
  labs(title = "CitiBike Trips per Day by User Type",
       x = "Day",
       y = "Count",
       color = " ",
       caption = "https://www.citibikenyc.com/system-data") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The time-series shows more variability in the number of rides per day for customers. This makes sense, as someone who bought a membership to the service seems more likely to consistently use it. Customers may come and go based on holidays, weekends or other factors.

d.
```{r, warning=F, message=F}
library(ggseas)

ggplot(trips_per_day_usertype, aes(x = start_date, y = n_trips, color = usertype)) + 
  geom_line() + 
  stat_rollapplyr(width = 7, align = "right", size = 1.5) +
  scale_x_date(date_breaks = "1 week", date_labels = "%B %d") +
  labs(title = "CitiBike Trips per Day by User Type with Rolling Average",
       subtitle = "Average Trips per 7 Days Consistent for Subscribers", 
       x = "Day",
       y = "Count",
       color = " ",
       caption = "https://www.citibikenyc.com/system-data") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

There is more fluctuation in the moving average of customers than in that of the subscribers.

***
***


#  Problem 3

(20 points)

**Understanding Autocorrelations and Autocorrelation Plots**

a. *(5 points)  Autocorrelations are just the correlation of a time series with itself at different lags.  Let's start by looking at a totally random, made up time series:  `rand_ts <- rnorm(1000)`.  This is just a bunch of random numbers from a Normal(0, 1) distribution.  Use `acf(rand_ts, plot = FALSE)` to look at the correlation of the "time series" and itself at different lags.  Since this "time series" is just 1000 random draws from a Normal(0, 1) distribution, most of the autocorrelations should be near zero.*

+  *Is this true?*  
+  *Which autocorrelation is NOT near zero?*  
+  *Does this make sense?  Why or why not?*

b. *(5 points) Autocorrelation plots show the correlation between the time series and itself at all of the different possible lags.  The (auto)correlations are plotted on the y-axis, and the lags are shown on the x-axis.  Create an autocorrelation plot of the random time series from (a), using base R graphics.  To do this, just type `acf(rand_ts, plot = TRUE)`.  In the resulting plot, the blue dotted lines indicate statistical significance.  Any correlation above the positive blue line or below the negative blue line is significantly different from zero.*

+  *Are any of the autocorrelations significantly different from zero in your plot?*  
+  *Does this make sense?  Why or why not?*


c. *(10 points)  Create two autocorrelation plots of the `trips_per_day_usertype` time series:  One for subscribers, and one for customers. First, we create the two autocorrelation time series.*

```{r, warning = FALSE, message = FALSE}
#bike_customer_acf <- acf(filter(trips_per_day_usertype,
#                                usertype == "Customer")$n_trips,
#                         plot = FALSE)
#bike_subsciber_acf <- acf(filter(trips_per_day_usertype,
#                                 usertype == "Subscriber")$n_trips,
#                          plot = FALSE)
```

*And then we generate the both of them using ggplot. Here below we only generate the plot for the customer auto correlation function.*

```{r}
#bike_customer_acf_df <- with(bike_customer_acf, data.frame(lag,acf))
#ggplot(bike_customer_acf_df, aes(x = lag, y= acf)) + geom_col()
```

*Generate the autocorrelation plot for the subscribers as well.*

*Add to **both** plots*:

+ *An appropriate title*
+ *An appropriate source*
+ *Appropriate axis labels*

*In both time series, you should see that the ACF has its greatest peak at `lag=7`. How do you interpret this, considering that we computed the ACFs on daily-level data? Does it make sense, given the context of the data (number of rental bike trips per day)?*

***
***


#  Problem 4 (optional, ungraded)

(0 points)

This is an optional practice problem, designed to teach you how to (1) use `igraph` to compute simple statistics on a network dataset, (2) make plots with the `ggnetwork` package instead of `ggraph`, and (3) graph a directed network instead of Problem 1's undirected network.

We encourage you to do this practice problem, especially if you are using a network dataset for your poster project, but no credit will be given for submitting it.

**Directed networks, `igraph`, and `ggnetwork`**

The `igraph` package makes network analysis easy!  Install and load the `igraph` and `igraphdata` packages.  Load the `UKfaculty` network into `R`:

```{r, warning = FALSE, message = FALSE}
#install.packages("igraph")
#install.packages("igraphdata")
library(igraph)
library(igraphdata)
data(UKfaculty)
```


a. Try out some of the built-in functions in the `igraph` package in order to summarize the UK faculty network:
+ How many nodes ("vertices") are in the network?  (Hint: use `vcount()`.)  
+ How many edges ("links") are in the network?  (Hint: use `ecount()`.)

b. Each of the node of this network is the personal friendship network of a faculty of a UK university. In order to get a vector of the node labels, you can use the function `V()`. Save this vector as `graph_node_label`.

c. The `igraph` package has some built-in functions for analyzing both the in-degree and out-degree of the nodes in the network. Use the `igraph` function `degree` to get:

+ The in-degrees of all nodes. Save them as `in_degree_count`. (Hint: set the `mode` argument as `'in'` in the `degree` function)
+ The out-degrees of all nodes. Save them as `out_degree_count`. (Hint: set the `mode` argument as `'out'` in the `degree` function)

Partial code is provided below.

```{r}
in_degree_count <- degree(UKfaculty, mode = "in")

# Also compute out_degree_count here
```


d. Create now a dataframe with the three vectors created above, containing the nodes labels, in-degrees, and out-degrees respectively. (You may need to `cbind()` the vectors together, then save the result as a data frame.) Create a scatterplot of the in-degrees versus the out-degrees and comment on whether there seems to be any relationship between the two.

e. Let's actually visualize the network itself!  Install and load the `ggnetwork` package.  This package is decently new (you can read more about how to use it [in this article on the package](http://www.r-bloggers.com/ggnetwork-network-geometries-for-ggplot2/).

You'll also need to install and load the `intergraph` package in order to use `igraph` objects with `ggnetwork`.  Finally, depending on which version of `ggplot2`, `R`, and RStudio you're running, you may need to update your packages to the latest versions that are on GitHub (see the first commented lines of code below).

Visualize the UK faculty network using the code below.
Be sure to:

+ Add a title
+ Add a caption 
+ Remove the x- and y-axis labels
+ Remove the x- and y-axis tick marks
+ Adjust the `color` legend
+ Adjust the `size` legend

```{r, warning = FALSE, message = FALSE}
## Run these only if install.packages() does not work for you
#devtools::install_github("briatte/ggnetwork")
#devtools::install_github("mbojan/intergraph")

library(ggnetwork)
library(intergraph)

uk_data <- fortify(UKfaculty)
uk_in_degrees <- in_degree_count[match(uk_data$vertex.names,
                                       1:length(in_degree_count))]
uk_data$in_degree <- uk_in_degrees
ggplot(uk_data, 
       aes(x, y, xend = xend, yend = yend)) +
  geom_edges(arrow = arrow(length = unit(0.3, "lines")), 
             aes(color = as.factor(Group)), alpha = 0.5) +
  geom_nodetext(aes(label = vertex.names, size = in_degree * 1.5), 
                color = "blue", fontface = "bold")
```

f. Look at the network plot above:

+ How many groups ("cliques") do there appear to be?  
+ What does the size of the nodes correspond to?  
+ What does the color of the edges correspond to?  
+ Is the graph a directed or an undirected graph?

***
***

