---
title: "Static Graphics Project"
author: "Julie Kim"
date: "4/6/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Read-In
===

1. Load necessary packages
2. Read in `tripdata.csv`, save to variable `trip.data`
3. Remove rows of `trip.data` that do not contain user information (`member_birth_year`, `member_gender`)
4. Check dimensions and first several rows of our new dataframe

```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(ggmap)
library(geosphere)
library(gridExtra)
library(lubridate)

trip.data = read.csv("tripdata.csv")

trip.data = trip.data[which(!(trip.data$member_gender == "" | is.na(trip.data$member_gender))),]
trip.data = trip.data[which(!(trip.data$member_birth_year == "" | is.na(trip.data$member_birth_year))),]
dim(trip.data)
head(trip.data)
```


Set-Up for Graph 1
===

```{r, message = FALSE, warning = FALSE}
# Find the number of departures and arrivals at each station
departures = table(trip.data$start_station_id)
departures = departures[order(as.numeric(names(departures)))]
arrivals = table(trip.data$end_station_id)
arrivals = arrivals[order(as.numeric(names(arrivals)))]

# Build a vector of all station id's
station_id = sort(unique(c(trip.data$start_station_id, trip.data$end_station_id)))

# Build a data frame of station id's, longitudes, latitudes, number of departures, number of arrivals, and total traffic
station.data = as.data.frame(station_id)
station.data$latitude
station.data$longitude
station.data$departures
station.data$arrivals

for(i in 1:length(station_id)) {
  index = which(trip.data$start_station_id == station_id[i])[1]
  station.data$latitude[i] = trip.data$start_station_latitude[index]
  station.data$longitude[i] = trip.data$start_station_longitude[index]
  
  d.index = which(names(departures) == station_id[i])
  station.data$departures[i] = ifelse(length(d.index) > 0, departures[d.index], 0)
  
  a.index = which(names(arrivals) == station_id[i])
  station.data$arrivals[i] = ifelse(length(a.index) > 0, arrivals[a.index], 0)
}

station.data$traffic = station.data$departures + station.data$arrivals
```


Graph 1
===

```{r, message = FALSE, warning = FALSE}
# Create map base around the geological coordinates of the San Francisco Bay Area (37.778786, -122.350061)
map.base = get_map(location = c(lon = -122.350061, lat = 37.778786),
                    source = "google",
                    maptype = "roadmap")

# Create the map plot
# Points represent each station
# The points are sized by the amount of traffic that each station receives, with `traffic` representing the number of arrivals plus the number of departures
map.stations = ggmap(map.base) + 
  geom_point(data = station.data, alpha = 0.4,
             aes(x = longitude, y = latitude, size = traffic)) +
  scale_x_continuous(limits = c(-122.45, -122.2)) + 
  scale_y_continuous(limits = c(37.725, 37.9)) +
  labs(title = "Exhibit 1: GoBike Stations in the Bay Area",
       x = "", y = "", 
       size = "Total traffic") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(),
        axis.ticks = element_blank(), legend.position = "bottom")

map.stations
```


Set-Up for Graph 2
===

+ In this graph, we want to observe trip distances--is there an "optimal distance" for which this means of transportation is used? Does that differ by gender? By age? By user type (subscriber/casual)?

```{r, message = FALSE, warning = FALSE}
# Create a new column of `trip.data` that categorically assigns member age groups
# Assumption: Data is collected from 2018, and age is equal to 2018 - member birth year
trip.data$member_age = 2018 - trip.data$member_birth_year
trip.data$member_age_group[trip.data$member_age <= 24] = "Young Adult"
trip.data$member_age_group[trip.data$member_age > 24 & trip.data$member_age <= 45] = "Middle Adult"
trip.data$member_age_group[trip.data$member_age > 45 & trip.data$member_age <= 65] = "Middle Age"
trip.data$member_age_group[trip.data$member_age > 65] = "Senior"
```


Graph 2
===

```{r, message = FALSE, warning = FALSE, fig.height = 10}
x = as.matrix(trip.data[,c("start_station_longitude", "start_station_latitude")])
y = as.matrix(trip.data[,c("end_station_longitude", "end_station_latitude")])
trip.data$distance = distHaversine(x, y)

age.dens = ggplot(trip.data, aes(x = distance, fill = member_age_group)) +
  geom_density(alpha = 0.5, adjust = 2) + 
  scale_x_continuous(limits = c(0, 8000)) +
  labs(title = "",
       x = "", y = "Proportion of Trips",
       fill = "By Member Age Group") +
  scale_fill_manual(values = c("navyblue", "forestgreen", "yellow", "dodgerblue1")) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin = unit(c(-0.5, 1, -0.5, 1), "cm"))

gender.dens = ggplot(trip.data, aes(x = distance, fill = member_gender)) +
  geom_density(alpha = 0.5, adjust = 2) + 
  scale_x_continuous(limits = c(0, 8000)) +
  labs(title = "",
       x = "", y = "Proportion of Trips",
       fill = "By Member Gender") +
  scale_fill_manual(values = c("hotpink2", "dodgerblue2", "grey")) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin = unit(c(1, 1, -0.5, 1), "cm"))

type.dens = ggplot(trip.data, aes(x = distance, fill = user_type)) +
  geom_density(alpha = 0.5, adjust = 2) + 
  scale_x_continuous(limits = c(0, 8000)) +
  labs(title = "",
       x = "Route Distance (m)", y = "Proportion of Trips",
       fill = "By User Type") +
  scale_fill_manual(values = c("dodgerblue1", "yellowgreen")) +
  theme(plot.margin = unit(c(-0.5, 1, 1, 1), "cm"))

grid.arrange(age.dens, gender.dens, type.dens, top = "Exhibit 2: Trip Distance by Age Group, Gender, and User Type")
```


Set-Up for Graph 3
===

```{r, message = FALSE, warning = FALSE}
tcust = trip.data[trip.data$user_type == 'Customer', ]
tsub = trip.data[trip.data$user_type == 'Subscriber', ]
trip.data2 <- transform(trip.data, age = as.integer(format(Sys.Date(), "%Y")) - trip.data$member_birth_year)
tripe.data2 = trip.data2[!(is.na(trip.data2$member_gender) | trip.data2$member_gender==""), ]
trip.data2 <- transform(trip.data2, day = weekdays(as.Date(trip.data2$start_time)))
trip.data2 <- transform(trip.data2, hour = hour(trip.data2$start_time))
trip.data2$hour<- as.factor(trip.data2$hour)
trip.data2$day<- as.factor(trip.data2$day)
head(trip.data2)
```


Graph 3
===

```{r, message = FALSE, warning = FALSE}
ggplot(data = trip.data2, aes(x = age, fill = member_gender)) + 
      geom_histogram(aes(y = ..density..), bins = 25)  +  
      scale_x_continuous(limits = c(5, 80)) + 
      facet_wrap(~ user_type, ncol = 2) + 
      labs(title = "Exhibit 3: Age Distributions by Subscriber Type",
       x = "Age",
       y = "Density",
       fill = "Gender") + scale_fill_manual(values=c("royalblue4", "forestgreen", "#56B4E9"))
```


Set-Up for Graph 4
===

```{r, message = FALSE, warning = FALSE}
trip.data2 %>%
  group_by(day, hour) %>%
  summarize(total.rides = n(),
    avg.duration = mean(duration_sec)) -> day_hour_rides

day_hour_rides %>%
  ungroup() %>%
  select(hour, day, total.rides, avg.duration) %>%
  mutate(total_duration = total.rides * avg.duration, 
         hour = factor(hour, levels = (0:23))) %>%
  group_by(hour, day) %>%
  summarise(count.rides = sum(total.rides), total.duration = sum(total_duration)) -> df.1

head(df.1)
```


Graph 4
===

```{r, message = FALSE, warning = FALSE}
ggplot(data = df.1, aes(x = day, y = hour, fill = count.rides)) + 
  geom_tile() + 
  scale_fill_continuous(low = 'white', high = 'dark green') +
  labs(title = "Exhibit 4: Bike Usage Focused on Weekdays",
       x = "Day of Week",
       y = "Hour of Day",
       fill = "Number of Rides") 
```


Graph 5
===

```{r, message = FALSE, warning = FALSE}
bikes <- read_csv("tripdata.csv")
bikes %>% mutate(weekend = ifelse(wday(start_time) > 5, "Weekend", 
                                  "Weekday")) %>% 
  ggplot(aes(x = start_time, color = weekend)) +
  geom_freqpoly(binwidth = 3600*24*7) +
  scale_x_datetime(limits = c(min(bikes$start_time), 
                              ymd_hms("2017-12-31 23:59:59"))) +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Exhibit 5: Distribution of Bike Usage over Time",
       x = "Date", 
       y = "Rides per Week",
       color = "")
```


Graph 6
===

```{r, warning = FALSE, message = FALSE}
station.names = strsplit(bikes$start_station_name, " [(]")
station.names = lapply(station.names, "[[", 1)
station.names = unlist(station.names)
bikes$start_station_name = station.names

popular_stations <- 
  bikes %>% mutate(start_station_name = start_station_name %>% 
           stringr::str_replace_all(pattern = " \\(", 
                                    replacement = "\n\\(")) %>%
  group_by(start_station_name) %>% 
  count(sort = TRUE) %>% filter(n > 11500)
  
bikes %>% mutate(start_station_name = start_station_name %>% 
           stringr::str_replace_all(pattern = " \\(", 
                                    replacement = "\n\\(")) %>% 
  filter(start_station_name %in% popular_stations$start_station_name) %>% 
   mutate(start_station_name = start_station_name %>% factor) %>% 
  ggplot(aes(x = start_time, color = start_station_name)) +
  geom_freqpoly(bins = 27) +
  scale_x_datetime(limits = c(min(bikes$start_time), 
                              ymd_hms("2017-12-31 23:59:59"))) +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Exhibit 6: Five Most Popular Stations over Time", 
       x = "Date",
       y = "Rides per Week", 
       color = "Starting Station")
```


Set-Up for Graphs 7 & 8
===

```{r}
#protein = cut(protein, c(-Inf, median(protein), Inf), labels = c("Low", "High"))

trip.data2 %>%
  group_by(day, hour, age, member_gender) %>%
  summarize(total.rides = n(),
            avg.duration = mean(duration_sec)) -> day_age_rides

day_age_rides %>%
  ungroup() %>%
  select(hour, day, age, member_gender, total.rides, avg.duration) %>%
  mutate(total_duration = total.rides * avg.duration, 
         hour = factor(hour, levels = (0:23)),
         age_general = cut(age, c(0, 24, 45, 65, Inf), labels = c("Young Adult", "Middle Adult", "Middle Age", "Senior"))) %>%
  group_by(hour, day, age, age_general, member_gender) %>%
  summarise(count.rides = sum(total.rides), total.duration = sum(total_duration)) -> df.5
df.5
```

Graph 7
===

```{r}
ggplot(data = df.5, aes(x = hour, y = count.rides, col = age_general)) + 
  geom_jitter(alpha = 0.1) +
  geom_smooth(method = "loess", aes(x = as.integer(hour), y = count.rides)) +
  coord_flip() +
  scale_color_manual(values = c("goldenrod3", "green4", "blue3", "grey")) +
  labs(title = "Exhibit 7: Number of Rides Each Hour of the Day",
       subtitle = "By Age Group",
       x = "Hour",
       y = "Number of Rides",
       col = "Age Group")
```



Graph 8
===

```{r}
ggplot(data = df.5, aes(x = hour, y = count.rides, col = member_gender)) + 
  geom_jitter(alpha = .1) +
  geom_smooth(method = "loess", aes(x = as.integer(hour), y = count.rides)) +
  coord_flip() +
  scale_color_manual(values = c("hotpink2", "dodgerblue2", "grey")) +
  labs(title = "Exhibit 8: Number of Rides Each Hour of the Day",
       subtitle = "By Gender",
       x = "Hour",
       y = "Number of Rides",
       col = "Gender") 
```


INSERT ADDITONAL GRAPHS ABOVE THIS SECTION
===

Extraneous Work: San Francisco Specific Data
===

```{r, message = FALSE, warning = FALSE}
# Get all data with longitudes smaller than -122.35, so that we grab only the data relevant to San Francisco
sf.data = trip.data[which(trip.data$start_station_longitude < -122.35 |
                            trip.data$end_station_longitude < -122.35),]
sf.stations = station.data[which(station.data$longitude < -122.35),]

# Create map base around the geological coordinates of San Francisco (37.778786, -122.350061)
sf.map.base = get_map(location = c(lon = -122.350061, lat = 37.778786),
                    source = "google",
                    maptype = "roadmap")

# Create the map plot
# Points represent each station
# The points are sized by the amount of traffic that each station receives, with traffic equalling the number of arrivals plus the number of departures
sf.map.stations = ggmap(map.base) + 
  geom_point(data = sf.stations, alpha = 0.4,
             aes(x = longitude, y = latitude, size = traffic)) +
  scale_x_continuous(limits = c(-122.45, -122.375)) + 
  scale_y_continuous(limits = c(37.725, 37.825))

sf.map.stations
```


